PREPRINT -- Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery -- PREPRINT.

### 2024-03-09 docs/_tex/index.tex
laurent's comments

A total of 9 prediction maps corresponding to the 9 drone flights were
obtained. Each prediction map is associated with a probability map,
indicating the probability of the selected class for every pixel. The
low-altitude flight conducted in Gafanaha, Portugal, represents the site
with the highest complexity (Figure~\ref{fig-GafLow}). Among the 5
vegetation classes on which the model was trained, 4 were present on
this site. On this site, there is a mixture of Chlorophyceae and
Rhodophyceae over the seagrass meadow. This is also where
Bacillariophyceae is most abundant. Although the seagrass bed is solely
composed of \emph{Zostera noltei}, various colors can be observed: dark
green (indicating healthy beds with 100\% coverage) and whitish/brown
(indicating beds where leaves are dying). Regardless of the color, the
meadow is predicted as Magnoliopsida by the model.

@@

\caption{\label{fig-GafLow}RGB orthomosaic (Left) and Prediction (Right)
of the low altitude flight of Gafanaha, Portugal. The total extent of
this flight is 3000m² with a resolution of 8 mm per pixel. Background
colors means intertidal area (Light Green) and land area (Light Grey).
The zoom covers an area equivalent to a 10-meter Sentinel-2 pixel size.}

@@

The high-altitude flight over Gafanha covered nearly 1 km² in total
(Figure~\ref{fig-GafHigh}). A channel delineating a small island was
masked in the prediction map due to sunglint and misclassification
caused by the turbid shallow water. Most of the intertidal area has been
classified as Magnoliopsida by the model, even though there is
discoloration of seagrass blades. Only a few pixels have been classified
as Chlorophyceae at this scale. Furthermore, the area that was
classified as Bacillariophyceae in the low-altitude flight remains
mostly classified as such in the high-altitude flight, though some
pixels were classified as Magnoliopsida. Patches of Rhodophyceae were
correctly classified. In the nothern part of the scene, near the land
limit, patches of \emph{Spartina sp.} were misclassified, either as
Magnoliopsida or as Phaeophyceae.

@@

\caption{\label{fig-GafHigh}RGB orthomosaic (Left) and Prediction
(Right) of the high altitude flight of Gafanaha, Portugal. The total
extent of this flight is about 1 km² with a resolution of 80 mm per
pixel. Background colors means intertidal area (Light Green), land area
(Light Grey) and water (Light Blue). The red triangle shows the extent
of the low altitude flight of Gafanha. The zoom covers an area
equivalent to a 10-meter Sentinel-2 pixel size.}

@@

The high altitude flight acquired over the inner lagoon of the Ria de
Aveiro is the largest of all flights, covering almost 1.5 km²
(Figure~\ref{fig-Boat}). On this site, only seagrass and red algae where
seen on the field. The classification provided consistent results, with
a patchy Magnoliopsida meadow mixed with Rhodophyceae on the eastern
part of the scene. As shown in the zoom (Figure~\ref{fig-Boat}), the
edges of the meadow can be colonised by Chlorophyceae.

@@

\caption{\label{fig-Boat}RGB orthomosaic (Top) and Prediction (Bottom)
of the flight made in the inner part of Ria de Aveiro Lagoon, Portugal.
The total extent of this flight is about 1.5 km² with a resolution of 80
mm per pixel. Background colors means intertidal area (Light Green),
land area (Light Grey) and water (Light Blue). The zoom covers an area
equivalent to a 10-meter Sentinel-2 pixel size.}

@@

The flight over L'Epine in Noirmoutier Island, France
(Figure~\ref{fig-Dike}) was conducted near a dike crossing the northern
part of the scene from west to east. Alongside this dike, brown algae
attached to rocks and stranded green algae could be found. Despite the
high mixture between Chlorophyceae and Magnoliopsida these two classes
were correctly discriminated by the classifier.

@@

\caption{\label{fig-Dike}RGB orthomosaic (Top) and Prediction (Bottom)
of Northern part of Noirmoutier Island, France. The total extent of this
flight is about 28 000 m² with a resolution of 80 mm per pixel.
Background colors means intertidal area (Light Green) and land area
(Light Grey). The zoom covers an area equivalent to a 10-meter
Sentinel-2 pixel size.}

@@

A total of 536,000 pixels was used to validate the Neural Network
classifier. The sites with the lowest and highest number of validation
data were Kerdrean (5557 pixels) and Marinha Lanzarote (159713 pixels),
respectively. Model global accuracy was 94.26\% with a Kappa coefficient
of 0.92 (Figure~\ref{fig-Validation}). The least performing site was
Gafanha High (global accuracy of 75.45\%) whereas Mataducos was the site
with the most accurate prediction (global accuracy of 98.05\%). Overall,
the classes Phaeophyceae, Magnoliopsida, Sediment and Rhodophyceae were
correctly classified with a balanced accuracy of 1, 0.96, 0.96 and 0.91
respectively. Bacillariophyceae was the least performing class (accuracy
of 0.72 ) mainly due to the confusion between Magnoliopsida and
Sediment.

@@

\caption{\label{fig-VIP}Variable Importance of the Neural Network
Classifier for each vegetation class. The bigger the slice, the more
important is the variable to predict accuratly this class. The bottom
right plot shows the drone standardised reflectance spectra of each
class. Each slice represent the VIP of both RAW and standardised
reflectance combined.}

@@

The computation of the variable importance makes it possible to identify
which wavelengths are the most useful for class prediction
(Figure~\ref{fig-VIP}). The spectral bands at 444 , 717 and 842 nm of
the Micasense camera are important for none of the vegetation classes.
The band at 531 nm is the only important predictor for the classifier to
accurately predict chlorophyceae. In fact, at this wavelength, the
Chlorophyceae spectra has the maximum reflectance among of all
vegetation classes. The bands at 531 and 740 nm are the most important
predictors for Phaeophyceae, corresponding to the minimum reflectance
among all classes. Bands at 475 and 560 nm are the most important
predictors for Bacillariophyceae and Rhodophyceae, respectively. Four
predictors, ranging from the green (560 nm) to the RedEdge (705 nm)
bands are important to accurately predict magnoliopsida.

@@

\caption{\label{fig-upscaling}Kernel density plot showing the proportion
of pixel well classified based on the percent cover of the class in high
altitude flight pixels of Gafanha, Portugal. Each subplot shows all the
pixels of the same classes on the hight altitude flight. Percent cover
of classes is retrieve using the result of the classification of the low
altitude flight of Gafanha, Portugal. The vertical dashed line shows the
0.85 probability of the model. Everything on the right of this line has
a probability higher then 0.85 and everuthing on the left of this line
has a probability lower.}

@@

Using the very high precision low altitude flight, we determined the
minimal percent cover required to classify correctly classify a given
class using the high altitude flight (Figure~\ref{fig-upscaling}). When
the percent cover of the class is 100 \%, big pixels are well classified
for all the classes excepted for Bare Sediment, where it's well
classified 80\% of the time. A vegetative percent cover of at least 80\%
is need to have all the big pixels well classified, at the exception of
Magnoliopsida that needs an higher percent cover (\textgreater90 \%) to
be well classified. Concerning the probability of each class, a really
high Percent cover is needed to confidently predict Bacillariophyceae.
To predict Chlorophyceae with a confidence of 0.85, a percent cover of
93 \% is needed, 90 \% for magnoliopsida, 92 \% for Rhodophyceae and 97
\% for Bacillariophyceae.

@@

\caption{\label{fig-Pigm}Photosynthetic and carotenoid pigments present
(Green) or absent (Red) in each taxonomic class present in the Neural
Network Classifier, along with their absorption wavelength measured with
spectroradiometer. Chl b: chlorophyll b, Chl c: chlorophyll c, Fuco:
fucoxanthin, Zea: zeaxanthin, Diato: diatoxanthin, Diadino:
diadinoxanthin, Neo: neoxanthin.}

@@

The primary objective of this study was to develop a method for the
accurate classification of macrophytes on intertidal mudflats,
specifically focusing on distinguishing between Chlorophyceae (green
algae) and marine Magnoliopsida (seagrasses) using multispectral drone
data. The ability to differentiate between various types of vegetation,
as demonstrated in this study, plays a critical role in ecological
monitoring and management. By distinguishing between seagrasses and
macroalgae, our approach facilitates targeted conservation strategies,
enabling more effective preservation and restoration efforts in coastal
ecosystems. The discrimination of seagrasses from green macroalgae
presents significant challenges, primarily due to the shared pigment
composition between these taxa, as depicted in Figure~\ref{fig-Pigm} but
also by the frequent spatial mixing of these green macrophytes. This
similarity in pigment composition, including photosynthetic and
carotenoid pigments like chlorophyll b and c, fucoxanthin, and
neoxanthin, results in spectral signatures that are often
indistinguishable to non exper eyes. This spectral overlap is further
complicated by the intermingling of seagrasses and green macroalgae in
the same spatial locations, posing a significant challenge for remote
sensing aimed at accurately mapping and monitoring coastal ecosystems.
Our study addresses these complexities by utilizing high-resolution
drone imagery with 10 spectral bands. \citep{Davies2023} has shown that
having at least 8 spectral bands ranging between 500 nm to 850 nm with
including a band at 530 nm and another one at 730 nm is crucial to
accurately discriminate green macrophytes.

@@

The Micasense RedEdge-MX DUAL camera used in this study meets these two
criteria, enabling the classifier to achieve 97\% of accuracy between
these two classes (Figure~\ref{fig-ValidationGreen}). Even if the
pigment compostion of green macrophytes is similar, differences in the
spectral shape can still be observed (Figure~\ref{fig-vegetation}).
Several factors can explain these differences such has different pigment
concentration or proportion, different cellular organisation and 3D
disposition of the plant a whole on the intertidal mudflat
\citep[\citep{kirk1994light},
\citep{hedley2018influence}]{beach1997vivo}.

The VIP analysis of the Neural Network model (Figure~\ref{fig-VIP})
shows that the 531 nm band is the most important spectral band for
accurately identifying Chlorophyceae. In fact, at this wavelength,
Chlorophyceae exhibits the highest reflectance among all other classes,
highlighting the difference in accessories carotenoid proportion between
seagrasses and green algae \citep{repolho2017seagrass}.

Concerning Phaeophyceae, the thick cell walls of plant of this class
make it really reflective in the infrared part of the spectra whereas
the presence of Fucoxanthin and Zeaxanthin result in a low reflectance
in the visible part of the electromagnetic spectra. These two key
features has been identified by the Neural Network as the two principal
predictors to accurately identify Brown algae (Figure~\ref{fig-VIP}).
Similarly, the presence of phycoerythrin and phycocyanin in Rhodophyceae
contributes to the lowest reflectance among all classes in the spectral
range of 560 to 615 nm (Figure~\ref{fig-VIP}). The band at 560 nm has
been identified as important for accurately identifying this class
Resolution impact on the prediction. Regarding Bacillariophyceae, the
VIP analysis (Figure~\ref{fig-VIP}) indicates that 475 nm is the most
important predictor for this class. This could be attributed to the
absence of lutein, neoxanthin, and zeaxanthin in diatoms, absorbing in
the blue part of spectra, resulting in the highest visible reflectance
among all classes. Furethemore, it is vegetation with the lowest
concentration of chlorphyll-a, pigment absorbing light both in the blue
and the red. The transparency of Bacillariophyceae makes the reflectance
of the sediment part of the overall reflectance of Bacillariophyceae,
further explaining the high reflectance in the blue.

@@

\caption{\label{fig-CompareRef}Comparison of reflectance retrieved from
both low-altitude and high-altitude flights over a common area. The red
dashed line represents a linear model wheras the black dashed line
represents a 1 to 1 relationship. Left (A) plots RAW data and right (B)
plots standardise data.}

@@

In this study, a key innovation lies in the utilization of two different
altitudes (12m and 120m) for constructing the neural network model. The
lower altitude flights (8mm spatial resolution) enable precise selection
of pure pixels representing the classes used in the neural network
model. This methodology implies a consistency between the reflectance of
both altitude. Figure~\ref{fig-CompareRef} (A) depicts the relationship
between reflectance from a low-altitude flight and a high-altitude
flight conducted over the same area. An apparent underestimation of
reflectance values is observed in the high-altitude flight, particularly
for higher values. Since both flights were conducted over vegetation
areas, the highest reflectance values correspond to the infrared part of
the spectrum.

The disparity in infrared reflectance may stem from temporal differences
between the flights, possibly resulting in a slightly drier intertidal
area and consequently higher infrared reflectance. This disparity poses
an issue for the methodology followed in the present study, relying
solely on one flight height for training. To address this issue, we
employed min/max standardized reflectance spectra as predictors for the
model (Equation~\ref{eq-std} ; Figure~\ref{fig-CompareRef} (B) ;
\citep{Cao2017}). This approach allows us to eliminate differences in
the level of reflectance, focusing instead on the shape of the spectra.
This is a key feature in building a model that can reliably predict
vegetation across geographical sites and seasons. It enables consistent
prediction of vegetation classes across variations in biomass and
variability in light conditions \citetext{\citealp[
]{fyfe2003spatial}; \citealp[
]{COSTA2021107018}; \citealp{piaser2023impact}}.

The Figure~\ref{fig-upscaling} demonstrates that a seagrass cover
percentage of 90 \% is necessary for confident prediction of seagrass
presence. This highlights a limitation of the methodology used to
construct the training dataset for the model. The dataset was composed
exclusively of pure pixels, which has resulted in the model's reduced
confidence when faced with lower percentages of seagrass cover.
Intertidal seagrasses exhibit strong phenology, with varying pigment
composition throughout the year\citetext{\citealp[
]{bargain2013seasonal}; \citealp{legare2022remote}}. This suggests that
outside of the biomass peak of seagrasses, this model may be less
accurate. Further investigation is warranted to explore this aspect.

@@

Climate change, global warming, alien and invasive species development,
coastal erosion, sealevel rise, while further impact coastal ecosystems
in the future \citetext{\citealp{SCHIBALSKI2022101414}; \citealp[
]{holon2018predictive}; \citealp{marquet2024global}}. The demand for
meaningful and efficient monitoring methods for these habitats has
therefore never been greater \citetext{\citealp[
]{muller2018satellite}; \citealp[
]{villalobos2023remote}; \citealp{oiry2021using}}. Our findings,
particularly the improved discrimination of vegetation types, underscore
the potential of drone-based remote sensing to support diverse
applications, from biodiversity conservation to climate change
adaptation strategies\textbf{.} Employing traditional sampling methods
to monitor these ecosystems is time and resource-intensive, and the
findings are often difficult to scale up. Earth observation methods can
bridge this gap and meet the needs for monitoring coastal ecosystems.
\citep{papathanasopoulou2019satellite}. The retrieval of Essential
Biodiversity Variables (EBVs) and Essential Ocean Variables (EOVs)
through satellite observations is increasingly common, enabling
comprehensive monitoring of entire ecosystems over extended time periods
\citetext{\citealp[
]{ratnarajah2023monitoring}; \citealp{ZOFFOLI2020112020}}. The
significance of monitoring waters, particularly in light of the Water
Framework Directive (WFD ; \citep{WFD2000}), cannot be overstated in
environmental management and conservation efforts. The WFD mandates the
achievement and maintenance of ``good ecological status'' for all
European waters, which necessitates a comprehensive understanding and
monitoring of aquatic ecosystems, including coastal habitats like
seagrass \citetext{\citealp[ ]{foden2007angiosperms}; \citealp[
]{nordlund2024one}; \citealp{Zoffoli2021}}.

Effective and efficient monitoring tools are essential for identifying
the impacts of human activities and natural changes on these ecosystems.
Drones, with their capability for high-resolution, multispectral
imagery, provide a novel and powerful means to rapidly and accurately
acquire ground truth data (i.e.~output of the model used in the present
study). This data is critical for calibrating and validating satellite
remote sensing applications, thereby enhancing our capacity to monitor
vast coastal areas systematically under the WFD framework. The
integration of drone technology facilitates a scalable approach to
environmental surveillance, offering significant advancements in the
spatial and temporal resolution of data collection. This, in turn,
supports the directive's objectives by enabling more informed and timely
management decisions for the preservation and restoration of aquatic
ecosystems.

@@

The utilization of drone-based remote sensing coupled with machine
learning techniques has proven to be an effective method for the
discrimination of seagrasses and macroalgae in coastal ecosystems. This
study has highlighted the potential for these technologies to enhance
the accuracy and efficiency of habitat monitoring, which is essential
for environmental management and conservation strategies. Through the
application of high-resolution multispectral imagery, the research
contributes to the broader understanding and surveillance of aquatic
ecosystems, aligning with objectives outlined in policies such as the
Water Framework Directive. The findings underscore the importance of
adopting advanced remote sensing tools in ecological studies and
environmental monitoring, providing a foundation for future research and
policy implementation aimed at ecosystem preservation and restoration.

>>>>

A total of nine prediction maps corresponding to the nine drone flights
were obtained. Each prediction map is associated with a probability map,
indicating the probability of the selected class for every pixel. The
low-altitude flight conducted in Gafanha, Portugal, represents the site
with the highest complexity (Figure~\ref{fig-GafLow}). Among the five
vegetation classes on which the model was trained, four were present on
this site, with Chlorophyceae and Rhodophyceae mixed with the seagrass
meadow. There was also Bacillariophyceae forming biofilms on parts of
sediment surface. Although the seagrass bed was solely composed of
\emph{Zostera noltei}, various colors of this could be observed from
dark green (corresponding to healthy leaves) to whitish/brown (when
leaves were discolored having lost their pigmentation). Regardless of
color, the meadow was predicted as Magnoliopsida by the model.

@@

\caption{\label{fig-GafLow}RGB orthomosaic (Left) and Prediction (Right)
of the low altitude flight of Gafanha, Portugal. The total extent of
this flight is 3000m² with a resolution of 8 mm per pixel. Background
colors indicate intertidal area (Light Green) and land area (Light
Grey). The zoom covers an area equivalent to a 10-meter Sentinel-2 pixel
size.}

@@

The high-altitude flight over Gafanha covered nearly 1 km² in total
(Figure~\ref{fig-GafHigh}). A channel delineating a small island was
masked in the prediction map due to sunglint and misclassification
caused by the turbid shallow water. Most of the intertidal area was
classified as Magnoliopsida by the model, including seagrass patches
with discolored leaves. Only a few pixels were classified as
Chlorophyceae at this scale. Furthermore, the area that was classified
as Bacillariophyceae in the low-altitude flight remained mostly
classified as such in the high-altitude flight, though some pixels were
classified as Magnoliopsida. Patches of Rhodophyceae were correctly
classified. In the nothern part of the scene, near the land limit,
patches of the schorre angiosperm \emph{Spartina sp.} were
misclassified, either as Magnoliopsida or as Phaeophyceae.

@@

\caption{\label{fig-GafHigh}RGB orthomosaic (Left) and Prediction
(Right) of the high altitude flight of Gafanha, Portugal. The total
extent of this flight is about 1 km² with a resolution of 80 mm per
pixel. Background colors indicate intertidal area (Light Green), land
area (Light Grey) and water (Light Blue). The red triangle shows the
extent of the low altitude flight of Gafanha presented in
Figure~\ref{fig-GafLow}. The zoom covers an area equivalent to a
10-meter Sentinel-2 pixel size.}

@@

The high altitude flight acquired over the inner lagoon of the Ria de
Aveiro was the largest of all flights, covering almost 1.5 km²
(Figure~\ref{fig-Boat}). The site was dominated by seagrass and the red
macroalgae \emph{Gracilaria vermicullophyla}. The classification
provided consistent results, with a patchy seagrass meadow mixed with
red macroalgae on the eastern part of the scene. As shown in the zoom
(Figure~\ref{fig-Boat}), the edges of the meadow were colonised by green
macroalgae

@@

\caption{\label{fig-Boat}RGB orthomosaic (Top) and Prediction (Bottom)
of the flight made in the inner part of Ria de Aveiro Lagoon, Portugal.
The total extent of this flight is about 1.5 km² with a resolution of 80
mm per pixel. Background colors indicates intertidal area (Light Green),
land area (Light Grey) and water (Light Blue). Each cover an area
equivalent to a 10-meter Sentinel-2 pixel size.}

@@

The flight over L'Epine in Noirmoutier Island, France
(Figure~\ref{fig-Dike}) was conducted near a dike crossing the northern
part of the scene from west to east. Alongside this dike, brown algae
for the order Fucales (\emph{Fucus spp.}, \emph{Ascophyllum nodosum})
were attached to sparse rocks and stranded green algae (\emph{ulva
spp.}) could be observed Despite the high mixture between Chlorophyceae
and Magnoliopsida these two classes were correctly discriminated by the
classifier.

@@

\caption{\label{fig-Dike}RGB orthomosaic (Top) and Prediction (Bottom)
of the Northern part of Noirmoutier Island, France. The total extent of
this flight is about 28 000 m² with a resolution of 80 mm per pixel.
Background colors indicate intertidal area (Light Green) and land area
(Light Grey). The zoom covers an area equivalent to a 10-meter
Sentinel-2 pixel size.}

@@

A total of 536,000 pixels was used to validate the Neural Network
classifier. The sites with the lowest and highest number of validation
data were Kerdrean (5557 pixels) and Marinha Lanzarote (159713 pixels),
respectively. Model global accuracy was 94.26\% with a Kappa coefficient
of 0.92 (Figure~\ref{fig-Validation}). The least performing site was
Gafanha High (global accuracy of 75.45\%) whereas Mataducos was the site
with the most accurate prediction (global accuracy of 98.05\%). Overall,
the classes Phaeophyceae, Magnoliopsida, Sediment and Rhodophyceae were
correctly classified with a balanced accuracy of 1, 0.96, 0.96 and 0.91
respectively. Bacillariophyceae was the least performing class (accuracy
of 0.72 ) mainly due to the confusion with Magnoliopsida and Sediment.

@@

\caption{\label{fig-VIP}Variable Importance of the Neural Network
Classifier for each taxonomic class. The bigger the slice, the more
important is the variable to predict accurately this class. The right
plot shows the drone raw and standardised reflectance spectra of each
class. Each slice represent the VIP of both raw and standardised
reflectance combined.}

@@

The computation of the variable importance made it possible to identify
which wavelengths were the most useful for class prediction
(Figure~\ref{fig-VIP}). The spectral bands at 444, 717 and 842 nm of the
Micasense camera did not came out as important for any of the vegetation
classes. The band at 531 nm was the only important predictor for the
classifier to accurately predict Chlorophyceae. In fact, at this
wavelength, the Chlorophyceae spectra have the highest reflectance among
of all vegetation classes. The bands at 531 and 740 nm were the most
important predictors for Phaeophyceae, corresponding to the lowest
reflectance among all classes. Bands at 475 and 560 nm are the most
important predictors for Bacillariophyceae and Rhodophyceae,
respectively. Four predictors, ranging from the green (560 nm) to the
RedEdge (705 nm) bands were important to accurately predict
Magnoliopsida.

@@

\caption{\label{fig-upscaling}Kernel density plot showing the proportion
of pixel well classified based on the percent cover of the class in high
altitude flight pixels of Gafanha, Portugal. Each subplot shows all the
pixels of the same classes on the hight altitude flight. Percent cover
of classes was retrieved using the result of the classification of the
low altitude flight of Gafanha, Portugal. The vertical dashed line shows
the 0.85 probability of the model. Everything on the right of this line
has a probability higher then 0.85 and everything on the left of this
line has a lower probability.}

@@

Using the very high resolution low altitude flight, we determined the
minimal percent cover required to classify correctly a given class using
the high altitude flight (Figure~\ref{fig-upscaling}). When the percent
cover of the class was 100 \%, coarser high-flight pixels are well
classified for all the classes except for Bare Sediment, which was well
classified 80\% of the time. This phenomenon is likely attributed to the
time gap between the two flights, allowing for microphytobenthos
migration, consequently altering the model's classification from bare
sediment to Bacillariophyceae. A percent cover of at least 80\% was
needed to have all the low resolution pixels well classified, at the
exception of Magnoliopsida that needs a higher percent cover
(\textgreater90 \%) to be well classified. Concerning the probability of
each class, the highest percent cover was needed to confidently predict
Bacillariophyceae. To predict Chlorophyceae with a confidence of 0.85, a
percent cover of 93 \% was needed, 90 \% for Magnoliopsida, 92 \% for
Rhodophyceae and 97 \% for Bacillariophyceae.

@@

\caption{\label{fig-Pigm}Photosynthetic and carotenoid pigments present
(Green) or absent (Red) in each taxonomic class dected by the Neural
Network Classifier, along with their absorption wavelength measured with
spectroradiometer. Chl b: chlorophyll b, Chl c: chlorophyll c, Fuco:
fucoxanthin, Zea: zeaxanthin, Diato: diatoxanthin, Diadino:
diadinoxanthin, Neo: neoxanthin., PE: phycoerythrin, PC: phycocyanin.}

@@

The primary objective of this study was to develop a method for the
accurate classification of macrophytes on intertidal mudflats,
specifically focusing on distinguishing between Chlorophyceae (green
macroalgae) and marine Magnoliopsida (seagrasses) using multispectral
drone data. The ability to differentiate between various types of
vegetation, as demonstrated in this study, plays a critical role in
ecological monitoring and management. By distinguishing between
seagrasses and macroalgae, our approach facilitates targeted
conservation strategies, enabling more effective preservation and
restoration efforts in coastal ecosystems. The discrimination of
seagrasses from green macroalgae presents significant challenges
\citetext{\citealp[ ]{oiry2021using}; \citealp{Bannari2022}}, primarily
due to the shared pigment composition between these taxa
\citep{mckenzie2020global}, as depicted in Figure~\ref{fig-Pigm} but
also by the frequent spatial mixing of these green macrophytes. These
green macrophytes have in common photosynthtic pigmetns, chlorophyll a
and b, and accessory carotenoid pigments zeaxantin, lutein and
neoxanthin. As a result the spectral signatures are often
indistinguishable in the visible wavelength region. This spectral
overlap is further complicated by the intermingling of seagrasses and
green macroalgae in the same spatial locations, posing a significant
challenge for remote sensing aimed at accurately mapping and monitoring
coastal ecosystems. Our study addresses these complexities by using
high-resolution drone imagery with 10 spectral bands. \citep{Davies2023}
have shown that having at least eight spectral bands ranging between 500
nm to 850 nm including a green band at 530 nm and another one in the
RedEdge region at 730 nm was crucial to accurately discriminate green
macrophytes.

@@

The Micasense RedEdge-MX DUAL camera used in this study meets these two
criteria, enabling the classifier to achieve 97\% of accuracy between
these two classes (Figure~\ref{fig-ValidationGreen}). Even if the
pigment composition of green macrophytes is similar, differences in the
spectral shape can still be observed (Figure~\ref{fig-vegetation}).
Several factors can explain these differences such has different pigment
concentration and/or proportion \citep{bargain2013seasonal}, different
cellular organisation and the disposition of the plant on the sediment
surface \citetext{\citealp[ ]{beach1997vivo}; \citealp[
]{kirk1994light}; \citealp{hedley2018influence}}.

The VIP analysis of the Neural Network model (Figure~\ref{fig-VIP})
identified that the 531 nm band as the most important spectral band for
accurately identifying Chlorophyceae. In fact, at this wavelength,
Chlorophyceae exhibited the highest reflectance among all other classes,
highlighting the difference in accessories carotenoid proportion between
seagrasses and green macroalgae \citep{repolho2017seagrass}.

Concerning Phaeophyceae, the thick cell walls of these macroalgae
\citep{charrier2021growth} make it more reflective in the infrared part
of the spectra \citep{Slaton2001} whereas the presence of Fucoxanthin
and Zeaxanthin result in a low reflectance in the visible part of the
electromagnetic spectra (Figure~\ref{fig-Pigm} ; Figure~\ref{fig-VIP}).
These two key features have been identified by the Neural Network as the
two principal predictors to accurately identify brown algae
(Figure~\ref{fig-VIP}). Similarly, the presence of phycoerythrin and
phycocyanin in Rhodophyceae contributes to the lowest reflectance among
all classes in the spectral range of 560 to 615 nm
(Figure~\ref{fig-VIP}). The band at 560 nm has been identified as
important for accurately identifying this class probably in relation to
the absorption by phycoerythrin (\textbf{REF}). Regarding
Bacillariophyceae, the VIP analysis (Figure~\ref{fig-VIP}) indicated
that 475 nm was the most important predictor for this class. classes.
This is not a usual diagnostic wavelength for Bacillariophyceae which is
mainly identified by the presence of fucoxanthin and chlorophyll c
(Table 11), but it correspond to the absorption of diadinoxanthin an
accessory pigment characteristic of this class
\citep{meleder2003spectrometric}. Furethemore, it is vegetation with the
lowest concentration of chlorphyll-a, pigment absorbing light both in
the blue and the red. The transparency of Bacillariophyceae makes the
reflectance of the sediment part of the overall reflectance of
Bacillariophyceae, further explaining the high reflectance in the blue.

@@

\caption{\label{fig-CompareRef}Comparison of reflectance retrieved from
both low-altitude and high-altitude flights over a common area. The red
dashed line represents a linear model whereas the black dashed line
represents a 1 to 1 relationship. Left (A) plots raw data and right (B)
plots standardised data.}

@@

In this study, a key innovation lies in the utilization of two different
altitudes (12m and 120m) for constructing the neural network model. The
lower altitude flights (8mm spatial resolution) enabled precise
selection of pure pixels representing the classes used in the neural
network model. This methodology implies a consistency between the
reflectance of both altitude. Figure~\ref{fig-CompareRef} depicts the
relationship between reflectance from a low-altitude flight and a
high-altitude flight conducted over the same area. Overall there was a
good agreement between the two altitudes. There was however a slight
underestimation for raw reflectance values in the high-altitude flight,
particularly for higher values (Figure~\ref{fig-CompareRef} A). Since
both flights were conducted over vegetation areas, the highest
reflectance values correspond to the infrared part of the spectrum.

The disparity in infrared reflectance may stem from temporal differences
between the flights, possibly resulting in a slightly drier intertidal
area and consequently higher infrared reflectance. This disparity poses
an issue for the methodology followed in the present study, relying
solely on one flight height for training. To address this issue, we
employed min/max standardized reflectance spectra as predictors for the
model (Equation~\ref{eq-std} ; \citep{Cao2017}). This approach allowed
us to eliminate the slight reflectance difference between the flights
(Figure~\ref{fig-CompareRef} B) and to focus on the shape of the spectra
in the visible part of the electromagnetic spectra, where different
pigmentation are associated to taxonomic dignostic features. This was a
key feature in building a model that could reliably predict vegetation
across geographical sites and seasons. It enabled consistent prediction
of vegetation classes across variations in biomass and variability in
light conditions \citetext{\citealp[ ]{fyfe2003spatial}; \citealp[
]{COSTA2021107018}; \citealp{piaser2023impact}}.

The Figure~\ref{fig-upscaling} demonstrates that a seagrass cover
percentage of 90 \% is necessary for confident prediction of seagrass
presence. This highlights a limitation of the methodology used to
construct the training dataset for the model. The dataset was composed
exclusively of pure pixels, which has resulted in the model's reduced
confidence when faced with lower percentages of seagrass cover.
Intertidal seagrasses exhibit marked phenology, with varying pigment
composition throughout the year\citetext{\citealp[
]{bargain2013seasonal}; \citealp{legare2022remote}}. Since the training
dataset has been made using well developed seagrass meadows, outside of
the biomass peak of seagrasses, this model may be less accurate. Further
investigation is warranted to explore this aspect.

@@

Climate change, global warming, alien and invasive species development,
coastal erosion, sealevel rise, while further impact coastal ecosystems
in the future \citetext{\citealp{SCHIBALSKI2022101414}; \citealp[
]{holon2018predictive}; \citealp{marquet2024global}}. The demand for
meaningful and efficient monitoring methods for these habitats has
therefore never been greater \citetext{\citealp[
]{muller2018satellite}; \citealp[
]{villalobos2023remote}; \citealp{oiry2021using}}. Our findings,
particularly the improved discrimination of seagrass from other
intertidal vegetation taxonomic classes, underscore the potential of
drone-based remote sensing to support diverse applications, from
biodiversity conservation to climate change adaptation strategies.
Employing traditional sampling methods to monitor these ecosystems is
time and resource-intensive, and the findings are often difficult to
scale up. Earth observation methods can bridge this gap and meet the
needs for monitoring coastal ecosystems.
\citep{papathanasopoulou2019satellite}. The retrieval of Essential
Biodiversity Variables (EBVs) and Essential Ocean Variables (EOVs)
through satellite observations is increasingly common, enabling
comprehensive monitoring of entire ecosystems over extended time periods
\citetext{\citealp[
]{ratnarajah2023monitoring}; \citealp{ZOFFOLI2020112020}}. The
significance of monitoring coastal environment, particularly in light of
the Water Framework Directive (WFD ; \citep{WFD2000}), cannot be
overstated in environmental management and conservation efforts. The WFD
mandates the achievement and maintenance of ``good ecological status''
for all European waters, which necessitates a comprehensive
understanding and monitoring of aquatic ecosystems, including coastal
habitats like seagrass \citetext{\citealp[
]{foden2007angiosperms}; \citealp[
]{nordlund2024one}; \citealp{Zoffoli2021}}.

Effective and efficient monitoring tools are essential for identifying
the impacts of human activities and natural changes on these ecosystems.
Drones, with their capability for high-resolution, multispectral
imagery, flight on demand, provide a novel and powerful means to rapidly
and accurately acquire ground truth data (i.e.~output of the model used
in the present study). These data are critical for calibrating and
validating satellite remote sensing applications, thereby enhancing our
capacity to monitor vast coastal areas systematically under the WFD
framework. A perspective of this work is to adapt the algorithm for
Sentinel-2 to see of the discrimination between seagrass and green
macroalgae is still operational at a coarser spatial resolution. The
integration of drone technology facilitates a scalable approach to
environmental surveillance, offering significant advancements in the
spatial and temporal resolution of data collection. This, in turn,
supports the directive's objectives by enabling more informed and timely
management decisions for the preservation and restoration of aquatic
ecosystems.

@@

The utilization of very high-resolution drone-based remote sensing
coupled with machine learning techniques has proven to be an effective
method for the discrimination of seagrasses from green macroalgae with a
multispectral resolution sensor. Standardized reflectance were
incorporated in the Neural Network model allowing for a better
discrimination of spectral features related to pigment absorption in the
visible. There was a striking difference for the variable of importance
to discriminate Magnoliopsida from Chlorophyceae. The latter was
essentially identify with the 451 nm spectral band while more spetral
bands were needed to identify the former, notably 650, 560, 668, and 705
nm. As the spectral bands of the Micasense RedEdge Dual MX are very
similar to those of Sentinel-2, we suggest that multispectral satellite
data have the potential to perform this discrimination between green
macrophytes. The algorithm should now be adapted to Sentinel-2 with all
the drone data from this study that can be used as training and
validation data. The findings underscore the importance of adopting
advanced remote sensing tools in ecological studies and environmental
monitoring, providing a foundation for future research and policy
implementation aimed at ecosystem preservation and restoration.
